### 2020-12-22 Outputs/Paper/paper.Rmd
merge

---
title: "2019 Canadian Federal Election"
subtitle: "....."
author: "Yijie Zhao"
date: "22 December 2020"
abstract: |
  | ...abstract
  |
  | **Keywords:** 2019 Canadian Federal Election; Liberal; Turnout; Multilevel Regression with Post-stratification.
output: pdf_document
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

# Data

Also bills and their average (Figure \@ref(fig:billssssss)). (Notice how you can change the height and width so they don't take the whole page?)

## ces2019 data

why cps19_votechoice as response variable? why not cps19_v_likely,cps19_vote_unlikely,cps19_vote_lean,cps19_v_advance?
sample size
other explanatory choosing

# Model

\begin{equation}
Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}  (\#eq:bayes)
\end{equation}

Equation \@ref(eq:bayes) seems useful, eh?

Here's a dumb example of how to use some references: In paper we run our analysis in `R` [@citeR]. We also use the `tidyverse` which was written by @thereferencecanbewhatever If we were interested in baseball data then @citeLahman could be useful.

# Results

# Discussion

## First discussion point

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

# Appendix

Code and data are available at: LINK.

\newpage

# References